Tabular Constraint Learning

### 2016-04-07 Paper/ECAI/tabular.tex
fixing the paper

A \textit{vector} is a subrange of a column or of a row that is type-consistent. If a vector is a subrange of a row (column), we say that it has a \textit{row} (\textit{column}) orientation. A \textit{group} is a subrange of vectors with the same orientation in a table. We use the following notation to refer to a row group $G$ in the $N$-th table with rows ranging from $a$ to $b$: $G = T_N[a{:}b,:]$, where $N,a,b$ are natural numbers. Similarly for a column group $G$ ranging from $a$ to $b$ columns in the $N$-th table we write $G = T_N[{:},a{:}b]$.

A \textit{constraint} is a triple of \textit{(\CName,\CSignature,\CFunction)}, where \CName is a textual name of the constraint; \CSignature is the signature of the constraint consisting of its arity and properties of the arguments, such as their types, sizes (properties might constrain sizes of two arguments as well, by for example, require them to be equal), etc; \CFunction is the function over arguments specified in \CSignature that represents constraint satisfaction. For example, a column-wise sum constraint has \CName equal to \textit{Y = SUM(X, col)}, its \CFunction sets arity to 2 and both arguments must be numeric and \sergey{Samuel, could you put the actual properties and \CFunction here?}

\subsection{Constraint learning algorithm description}
Let us describe the key steps and functions in Algorithm \ref{algo:tcl}. Essentially, the algorithm has two steps: candidate group generation and subgroup satisfaction search, which is in line with the well-known in AI paradigm of ``generate-and-test'' \cite{whaisasp}. Let us elaborate on each step in detail.

\paragraph{Candidate group generation} $\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups that are legitimate solution candidates e.g., $X,Y$ are variables of \textit{Y = SUM(X)} and $X,Y$ satisfy the following constraints \sergey{Samuel, can you add here, what is actually used for the sum now?}. Essentially, the group generation step is a constraint satisfaction problem associated with the specific constraint. However, many constraints have the same candidate generation procedures, e.g., min, max, avg, count, etc.

\paragraph{Subgroup satisfaction search} $\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function looking for the subsets of vectors in the candidates satisfying the constraint. If multiple subsets satisfy the constraint, a maximal is selected. If $G_1,G_2$, associated with the variables $X,Y$ in the constraint \textit{Y = SUM(X)}, are candidates, then \findassignment selects a single vector $y$ in $G_2$ and consecutive subset of vectors $x$ in $G_1$ such that the sum over $x$ is equal to $y$. For example, in Figure \ref{fig:main_example} the group $G = T_1[{:},3{:}7]$ can be used for both $X$ and $Y$ in the row-wise sum constraint, then as $x$ would be selected $T_1[{:},3{:}6]$ and as $y$ would be $T_1[{:},7]$.

\subsection{Constraints}
\sergey{we need to explain that constraints order is a partial order, which is a DAG}
\begin{table}
  \centering
  \begin{tabular}{lcc}
    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ 
  \end{tabular}
  \caption{Tabular Constraints}
  \label{table:constraints}
\end{table}

\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $D$ -- dataset, (optional: tables $T$, groups $G$)
    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
    \If{$T$ is \textbf{not} provided}
      \State $T \gets \extracttables(D)$
    \EndIf
    \If{$G$ is \textbf{not} provided}
      \State $G \gets \extractgroups(D, T)$
    \EndIf
    \State $S \gets \learnconstraints(G)$
    \State $S \gets \postprocess(S)$
    \State \Return $S$
\end{algorithmic}
\caption{Workflow}
\label{algo:workflow}
\end{algorithm}

>>>>

A \textit{vector} is either a column or a row that is type-consistent. If a vector is a row (column), we say that it has a \textit{row} (\textit{column}) orientation. A \textit{group} is a subrange of vectors with the same orientation in a table. We use the following notation to refer to a row (column) group $G$ in a table $T$ with rows (columns) ranging from $a$ to $b$: $G = T[a{:}b,:]$ ($G = T[{:},a{:}b]$), where $a,b$ are natural numbers.

An \textit{Excel constraint} is a triple \textit{(\CName, \CSignature, \CFunction)}, where \textit{\CName} is a textual name of the constraint; \textit{\CSignature} is a set of constraints specifying the properties of the arguments, such as their types, e.g., to require them to be integers or sizes, e.g. the length of vectors in the arguments must be equal, i.e., these are constraints on the group meta-information not on the actual group content; \textit{\CFunction} is a set of constraints specifying that the data in the subgroups satisfies the function. For example, an excel constraint \textit{rank} has \CName \textit{Y = RANK(X)}; its \CSignature is: $X$ is numeric, $Y$ is integer and the length of vectors in $X$ is the same as in $Y$; its \CFunction is the following constraint: a pair of vectors $x,y$ is a solution iff $x \in X, y \in Y$ and each value in $x$ has the rank (possibly with ties) specified in $y$.

\subsection{Constraint learning algorithm description}
Let us describe the key steps and functions in Algorithm \ref{algo:tcl}. Essentially, the algorithm has two steps: candidate group generation and subgroup satisfaction search, which is in line with the ``generate-and-test'' paradigm that is well-known in AI \cite{whaisasp}. Let us elaborate on each step in detail.

\paragraph{Candidate group generation} $\generategroups(\textit{Constraint,GroupSet,Solutions})$ is the function generating tuples of groups that are legitimate solution candidates e.g., $X,Y$ are variables of \textit{Y = RANK(X)} and $X,Y$ satisfy the constraints from \CSignature above. Essentially, the group generation step is a constraint satisfaction problem associated with the specific constraint. However, many constraints have the same candidate generation procedures, e.g., sum, min, max, avg, count, etc.

\paragraph{Subgroup satisfaction search} $\findassignment(\textit{Constraint,Candidates,Solutions})$ is the function looking for the subsets of vectors in the candidates satisfying the constraint. If multiple subsets satisfy the constraint, a maximal is selected. If $G_1,G_2$, associated with the variables $X,Y$ in the constraint \textit{Y = RANK(X)}, are candidates, then \findassignment selects a single vector $y$ in $G_2$ and a single vector $x$ in $G_1$ such that $x$ is ranked by $y$. For example, in Figure \ref{fig:main_example} the group $G = T_1[{:},3{:}8]$ serves as $X$ and $Y$ and $x$ can be picked as $T_1[{:},7]$ and $y$ would be $T_1[{:},8]$.

\subsection{Constraints}
The set of Excel constraints has a partial order in which they should be learned. For certain constraints this order does not matter, such as rank or product. For many others they have to be learned in the order, which is specified as a DAG in Figure \ref{fig:learning_order} \sergey{add the figure here}. This figure specifies a potential parallelization of the computation, since each connected component is independent of the others.

\begin{table}
  \centering
  \begin{tabular}{lcc}
    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\ 
  \end{tabular}
  \caption{Tabular Constraints}
  \label{table:constraints}
\end{table}

\subsection{Workflow}
\begin{algorithm}[thb]
  \begin{algorithmic}
    \footnotesize
    \State \textbf{Input:} $D$ -- dataset, (optional: tables $T$, groups $G$)
    \State \textbf{Output:} $S$ -- learned constraints with their satisfaction assignment
    \If{$T$ is \textbf{not} provided}
      \State $T \gets \extracttables(D)$
    \EndIf
    \If{$G$ is \textbf{not} provided}
      \State $G \gets \extractgroups(D, T)$
    \EndIf
    \State $S \gets \learnconstraints(G)$
    \State $S \gets \postprocess(S)$
    \State \Return $S$
\end{algorithmic}
\caption{Workflow}
\label{algo:workflow}
\end{algorithm}
